\Title

### 2020-11-10 manuscript/manuscript.tex
Improve manuscript

\subsection{Generalization of equivalent sources}

@@

Worth noting that solving equations~\ref{eq:linear-problem-g}
and~\ref{eq:linear-problem} is equivalent to performing an inversion and thus
solving an ill-posed problem by nature due to the non-uniqueness and the
instability of the possible solutions.
Therefore is usually needed to add regularization functions in order to obtain
stable solutions
(\citealp[p.~52]{menke1989}; \citealp{cooper2000}; \citealp{oliveira2013}).
Worth noting that the number of sources can either be equal, greater or lower
than the number of observation points, leading to well, under or overdetermined
linear problem, respectively.

% Add other possible applications to the EQL like RTP or upward continuation.

\subsection{Equivalent sources implementation}

The subsequent comparisons between strategies for interpolating data through
equivalent sources are carried out by the equivalent sources implementation in
Harmonica (v0.1.0) \citep{uieda2020}.

It solves equation~\ref{eq:linear-problem} through least squares with
a zeroth-order Tikhonov \citep{tikhonov1977} regularization in order to
stabilize the solution.
It consists in minimizing the misfit function:

@@

\noindent where $\left\lVert \cdot \right\rVert ^ 2$ symbolizes the L$_2$ norm
and $\lambda_d$ is a \emph{damping} parameter with the same units as the
Jacobian matrix elements.

In order to simplify the choice of this parameter we scale the Jacobian matrix
so each column has unit variance.
We define a diagonal matrix $\mathbf{S}$, whose elements corresponds to the
standard deviation of each column of the Jacobian matrix $\mathbf{A}$:

@@

\begin{equation}
    \hat{\mathbf{c}}_\text{est} =
        \left[
            \hat{\mathbf{A}}^\text{T} \hat{\mathbf{A}}
            + \lambda^2 \mathbf{I}
        \right]\inv
        \hat{\mathbf{A}}^\text{T}
        \mathbf{d}.
\end{equation}

@@

\begin{equation}
    \hat{\mathbf{c}}_\text{est} =
    \hat{\mathbf{A}}^\text{T}
        \left[
            \hat{\mathbf{A}} \hat{\mathbf{A}}^\text{T}
            + \lambda^2 \mathbf{I}
        \right]\inv
        \mathbf{d}.
\end{equation}

Once the scaled coefficients are obtained, the sources' coefficients
$\mathbf{c}_\text{est}$ can be obtained by rescaling
$\hat{\mathbf{c}}_\text{est}$:

\begin{equation}
    \mathbf{c}_\text{est} = \mathbf{S}\inv \hat{\mathbf{c}}_\text{est}.
\end{equation}

\section{Source distributions}

% Describe the proposed source distributions as combination of layouts and
% depths strategies.

@@

% Worth noting that the \emph{source below data} generates a number of sources
% equal to the data points, obtaining a well determined linear problem.
% The \emph{block averaged sources} layouts can generate a number of point
% sources equal or lower than the number of data points, well or
% overdetermined.
% The \emph{grid sources} is the only layout that can generate either a well,
% over or under determined problem.

@@

The combination of the three source layouts with the three strategies to define
the depth of the sources define seven different source distributions (see
Table~\ref{tab:source-distributions}).
The \emph{grid sources} is only compatible with the \emph{constant depth}
scheme.

@@

@@

Our goal is to to compare the performance of all source distributions from
Table~\ref{tab:source-distributions} when predicting harmonic field data using
the equivalent layer technique.
To do so we have created a synthetic model made out of prisms of different
shape and density contrasts.
The gravitational effect of this synthetic model have been computed on
a regular grid (\emph{target} grid) at a constant height and on two different
set of observation points that we will call \emph{synthetic surveys}: a ground
and an airborne survey.
The data from each synthetic survey is interpolated on the points of the
\emph{target} grid and then scored against the \emph{target} grid.
We can then quantitatively compare how accurate is each source distribution at
predicting unobserved data on each synthetic survey and quantify the
computational load needed to do it.

>>>>

\subsection{Generalized equivalent sources}

@@

\subsection{Fitting source coefficients}
\label{sec:eql-inversion}

Solving equations~\ref{eq:linear-problem-g} and~\ref{eq:linear-problem} is
equivalent to performing an inversion and thus solving an ill-posed problem by
nature due to the non-uniqueness and the instability of the possible solutions.
Therefore is usually needed to add regularization functions in order to obtain
stable solutions
(\citealp[p.~52]{menke1989}; \citealp{cooper2000}; \citealp{oliveira2013}).

We can obtain the values of the source coefficients $\mathbf{c}$ that better
reproduce the measured field by minimizing an objective or misfit function
$\phi(\mathbf{c})$ through a least squares method:

@@

\noindent where $\left\lVert \cdot \right\rVert ^ 2$ symbolizes the L$_2$ norm
and $\lambda_d$ is a \emph{damping} parameter with the same units as the
Jacobian matrix elements.
The second term of the equation applies a zeroth-order Tikhonov
\citep{tikhonov1977} regularization to stabilize the solution.

In order to simplify the choice of the \emph{damping} parameter we scale the
Jacobian matrix so each column has unit variance.
We define a diagonal matrix $\mathbf{S}$, whose elements corresponds to the
standard deviation of each column of the Jacobian matrix $\mathbf{A}$:

@@

\begin{equation}
    \hat{\mathbf{c}} =
        \left[
            \hat{\mathbf{A}}^\text{T} \hat{\mathbf{A}}
            + \lambda^2 \mathbf{I}
        \right]\inv
        \hat{\mathbf{A}}^\text{T}
        \mathbf{d}.
\end{equation}

@@

\begin{equation}
    \hat{\mathbf{c}} =
    \hat{\mathbf{A}}^\text{T}
        \left[
            \hat{\mathbf{A}} \hat{\mathbf{A}}^\text{T}
            + \lambda^2 \mathbf{I}
        \right]\inv
        \mathbf{d}.
\end{equation}

Once the scaled coefficients are obtained, the sources' coefficients
$\mathbf{c}$ can be obtained by rescaling
$\hat{\mathbf{c}}$:

\begin{equation}
    \mathbf{c} = \mathbf{S}\inv \hat{\mathbf{c}}.
\end{equation}

\subsection{Source distributions}

@@

@@

The combination of the three source layouts with the three strategies to define
the depth of the sources define seven different source distributions (see
Table~\ref{tab:source-distributions}).
The \emph{grid sources} is only compatible with the \emph{constant depth}
scheme if no further information is provided.

@@

\subsection{Gradient boosted equivalent sources}

Reducing the number of sources might help to lower the amount of computer
memory needed to fit their coefficients, although it may be not enough for
a very large set of observation points.
Even if we can greatly reduce the number of sources, having too many data
points would still produce a Jacobian matrix large enough that cannot fit in
memory.
This problem was already faced by other authors (CITE PEOPLE).
Here we propose an iterative algorithm for fitting the source coefficients
using a gradient boosting \citep{friedman2001, friedman2002} strategy.

Lets assume we have $N$ measurements of a harmonic field on observations points
$\{\mathbf{p}_i\}_{i=1}^{N}$ stored in the vector $\mathbf{d} = (d_1, d_2,
\ldots, d_N)^T$. We have defined a set of $M$ point sources located on
$\{\mathbf{q}_j\}_{j=1}^{M}$. Instead of trying to fit the entire set of
coefficients $\mathbf{c}$ at once, we can reduce the problem to fitting $K$
different sets of equivalent sources: we define $K$ overlapping windows of
equal size that cover the whole observation area, and then we create one set of
equivalent sources per window formed by the $\{\mathbf{q}_m\}_{m=1}^{M_k}$
sources that fall inside the $k$-th window.
The observed data should be equal to the superposition of the
effect of every set of equivalent sources:

\begin{equation}
    \mathbf{d} = \sum_{k=1}^K \alpha_k \mathbf{A}_k \mathbf{c}_k,
\end{equation}

\noindent where $\mathbf{c}_k$ is the vector of coefficients for the $k$-th
set of equivalent sources; $\mathbf{A}_k$ is the Jacobian matrix between
those sources and every data point; and each $\alpha_k$ is a constant called
\emph{step-size parameter}.

By applying gradient boosting we can fit each $\mathbf{c}_k$ vector
iteratively: on each iteration a residue vector $\mathbf{r}_k$ (initialized
with the observed data) is used to fit the $\mathbf{c}_k$
coefficients and $\alpha_k$ step-size parameter, and then updated by removing
the effect of the same sources, as follows:

\begin{equation}
    \mathbf{r}_0 = \mathbf{d},
\end{equation}

\begin{equation}
    \mathbf{r}_k = \mathbf{r}_{k-1} - \alpha_k \mathbf{A}_k\mathbf{c}_k.
    \label{eq:residue}
\end{equation}

\noindent On each iteration the $\mathbf{c}_k$ vector and the $\alpha_k$
parameter can be fitted so they minimize the misfit function:

\begin{equation}
    \phi(\mathbf{c}_k) =
        \left\lVert
            \mathbf{r}_{k-1} - \alpha_k \mathbf{A}_k\mathbf{c}_k
        \right\rVert ^ 2
        + \lambda_d \left\lVert \mathbf{c}_k \right\rVert ^2,
\end{equation}

\noindent following the same method described on Section
\ref{sec:eql-inversion}.
Because the $\mathbf{A}_k$ matrices have only $N \times M_k$ elements,
adjusting the size of the overlapping windows can help in reducing the computer
memory needed to fit the source coefficients and ultimately to interpolate the
observed data.
Nevertheless, in case we have too many data points that make the $\mathbf{A}_k$
matrices large enough that they cannot be stored in memory, we can apply one
more simplification on the fitting process: ignore the effect of the sources on
the $k$-th window on any observation point that falls outside it.

Assuming that there are $N_k$ observation points that fall under the $k$-th
window, we can fit the $\mathbf{c}_k$ coefficients by minimizing the following
misfit function:

\begin{equation}
    \phi(\mathbf{c}_k) =
        \left\lVert
            \tilde{\mathbf{r}}_{k-1} - \tilde{\mathbf{A}}_k\mathbf{c}_k
        \right\rVert ^ 2
        + \lambda_d \left\lVert \mathbf{c}_k \right\rVert ^2,
\end{equation}

\noindent where $\tilde{\mathbf{r}}_{k-1}$ is the residue vector containing
only the elements that correspond to the $\{\mathbf{p}_n\}_{n=1}^{N_k}$
observation points that fall under the $k$-th window; and
$\tilde{\mathbf{A}}_k$ is the Jacobian matrix between the observation points
and sources that fall under the same window.
Because the $\tilde{\mathbf{A}}_k$ matrices will have $N_k \times M_k$
elements, the computer memory needed for fitting the $\mathbf{c}_k$
coefficients is greatly reduced.

Once the $\mathbf{c}_k$ coefficients are fitted following the same method
described on Section \ref{sec:eql-inversion}, , we need to obtain the value
of the step-size parameter $\alpha_k$.
This can be achieved by minimizing the norm of the difference between the
previous residue $\mathbf{r}_{k-1}$ and the effect of the $k$-th set of
equivalent sources on every observation point:

\begin{equation}
    \alpha_k = \underset{\alpha}{\mathrm{argmin}}
    \left(
        \left\lVert
            \mathbf{r}_{k-1} - \alpha \mathbf{A}_k \mathbf{c}_k
        \right\rVert ^ 2
    \right),
\end{equation}

\noindent which leads to:

\begin{equation}
    \alpha_k = \frac{
        \mathbf{r}_{k-1}^T \mathbf{A}_k \mathbf{c}_k
    }{
        \left(\mathbf{A}_k \mathbf{c}_k\right)^T \mathbf{A}_k \mathbf{c}_k
    }.
\end{equation}

This step will improve the convergence of the iterative algorithm and also will
introduce at some extent the effect of the $k$-th sources on every data point,
which was neglected on the fitting of the $\mathbf{c}_k$ coefficients.

Once both the $\alpha_k$ parameter and the $\mathbf{c}_k$ coefficients are
fitted, equation \ref{eq:residue} is applied to update the residue: compute the
effect of the sources belonging to the $k$-th window on every observation
point. Worth noting that when calculating the effect of any number of sources
we don't need to pre-compute the entire Jacobian matrix: we can compute the
effect of the sources on each observation point by a simple addition of terms,
what means computing each Jacobian element at once when needed.

Because the number of windows is fixed before starting the iteration process,
its number of steps are also fixed to $K$. Moreover, the windows are randomly
shuffled before starting the process to prevent the prioritization of clustered
sets of equivalent sources.

After all $\alpha_k$ parameters and $\mathbf{c}_k$ coefficients vectors are
fitted, we can predict the effect of every equivalent source on any point as
follows:

\begin{equation}
    V(\mathbf{p}) =
    \sum\limits_{k=1}^K \sum\limits_{m=1}^{M_k}
    \frac{\alpha_k {c_k}_m}{|\mathbf{p} - \mathbf{q}_m|}.
\end{equation}

@@

Our first goal is to to compare the performance of all source distributions
from Table~\ref{tab:source-distributions} when predicting harmonic field data
using the equivalent layer technique.
To do so we have created a synthetic model made out of prisms of different
shape and density contrasts.
The gravitational effect of this synthetic model have been computed on
a regular grid (\emph{target} grid) at a constant height and on two different
set of observation points that we will call \emph{synthetic surveys}: a ground
and an airborne survey.
The data from each synthetic survey is interpolated on the points of the
\emph{target} grid and then scored against the \emph{target} grid.
We can then quantitatively compare how accurate is each source distribution at
predicting unobserved data on each synthetic survey and quantify the
computational load needed to do it.
